---
title: "Dose-Response and Dynamic Estimands"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Dose-Response and Dynamic Estimands}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(causaldosemc)
options(width = 100)
```

This vignette covers post-fit continuous dose-response estimation, prediction,
dynamic/pathwise summaries, simultaneous bootstrap bands, and bounded-bias
sensitivity analysis for downstream targets.

## Fit a parent baseline model

```{r}
panel <- simulate_cdmc_data(
  n_units = 16,
  n_times = 10,
  beta = 0.9,
  lag_beta = 0.3,
  noise_sd = 0.05,
  switch_on_prob = 0.17,
  switch_off_prob = 0.42,
  seed = 404
)

fit <- cdmc_fit(
  data = panel,
  outcome = "y",
  dose = "dose",
  unit = "unit",
  time = "time",
  covariates = "x1",
  lambda = 0.18,
  washout = 1,
  lag_order = 1,
  rank_max = 3,
  seed = 404
)

fit
```

## Fit a post-baseline dose-response model

```{r}
dose_response <- cdmc_dose_response(
  fit,
  model = "spline",
  lag_order = 1,
  df = 4
)

predict(dose_response, dose = seq(-1, 1, length.out = 5), type = "response")
```

For explicit histories, supply `dose_lag0`, `dose_lag1`, and any additional lag
columns up to the fitted order.

```{r}
history_grid <- data.frame(
  dose_lag0 = c(0, 0.5, 1),
  dose_lag1 = c(0, 0.25, 0.5)
)

predict(dose_response, history = history_grid, type = "response")
predict(dose_response, history = history_grid, type = "slope")
```

If `mgcv` is installed, a nonlinear GAM dose-response can be fit the same way.

```{r}
if (requireNamespace("mgcv", quietly = TRUE)) {
  dose_response_gam <- cdmc_dose_response(
    fit,
    model = "gam",
    lag_order = 1,
    df = 4
  )

  predict(dose_response_gam, history = history_grid, type = "response")
}
```

## Evaluate dynamic pathwise effects

```{r}
dynamic_paths <- cdmc_dynamic_estimand(
  dose_response,
  history = data.frame(
    dose_lag0 = c(1, 0.5),
    dose_lag1 = c(0.5, 1)
  ),
  reference_history = data.frame(
    dose_lag0 = c(0, 0),
    dose_lag1 = c(0, 0)
  ),
  type = "contrast",
  aggregate = "both",
  labels = c("front_loaded", "back_loaded")
)

dynamic_paths
dynamic_paths$estimate_table
```

## Bootstrap path summaries with simultaneous bands

```{r}
dynamic_boot <- cdmc_bootstrap(
  dynamic_paths,
  n_boot = 10,
  statistics = "estimate",
  simultaneous = TRUE,
  seed = 404
)

dynamic_boot$summary
```

## Add bounded-bias sensitivity summaries for predictions

```{r}
prediction_bounds <- cdmc_sensitivity_bounds(
  dose_response,
  statistics = "prediction",
  prediction_history = history_grid,
  prediction_type = "response",
  gamma_grid = c(0, 0.25, 0.5)
)

summary(prediction_bounds)
```

Linear and spline post-fit dose-response models support the exact stored-map
bounded-bias layer shown above. For nonlinear GAM, tree, or forest post-fit
models, use `perturbation_layer = "optimization"` to switch to the local refit
sensitivity layer.