---
title: "DR Weighting Workflows"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DR Weighting Workflows}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(causaldosemc)
options(width = 100)
```

This vignette covers the main weighting paths for `cdmc_dr_fit()`: internal
GPS weighting, balancing-based DR fitting, reusable external weights, and DR
replay under alternative nuisance choices.

To keep the focus on weighting rather than support trimming, the examples below
set `washout = 0`.

## Build a panel with richer continuous-dose variation

```{r}
panel <- simulate_cdmc_data(
  n_units = 24,
  n_times = 10,
  beta = 0.85,
  lag_beta = 0.2,
  noise_sd = 0.05,
  switch_on_prob = 0.12,
  switch_off_prob = 0.5,
  seed = 303
)

original_dose <- panel$dose
active <- abs(original_dose) > 0

panel$dose[active] <- sign(original_dose[active]) * (
  0.2 + 0.6 * panel$x1[active] + 0.1 * panel$time[active] / max(panel$time)
)
panel$y <- panel$y + 0.5 * (panel$dose - original_dose)

head(panel)
```

## Fit DR with an internal Gaussian GPS model

```{r}
dr_gps <- cdmc_dr_fit(
  data = panel,
  outcome = "y",
  dose = "dose",
  unit = "unit",
  time = "time",
  covariates = "x1",
  weight_method = "gaussian_gps",
  gps_model = "linear",
  n_folds = 2L,
  lambda = 0.18,
  washout = 0,
  lag_order = 0,
  rank_max = 3,
  seed = 303
)

dr_gps
```

If `mgcv` is installed, the same Gaussian GPS path can use a smoother mean
model for the assignment nuisance layer.

```{r}
if (requireNamespace("mgcv", quietly = TRUE)) {
  dr_gam <- cdmc_dr_fit(
    data = panel,
    outcome = "y",
    dose = "dose",
    unit = "unit",
    time = "time",
    covariates = "x1",
    weight_method = "gaussian_gps",
    gps_model = "gam",
    gps_df = 4L,
    n_folds = 2L,
    lambda = 0.18,
    washout = 0,
    lag_order = 0,
    rank_max = 3,
    seed = 303
  )

  dr_gam
}
```

## Fit DR with internal kernel-balance weights

```{r}
dr_balance <- cdmc_dr_fit(
  data = panel,
  outcome = "y",
  dose = "dose",
  unit = "unit",
  time = "time",
  covariates = "x1",
  weight_method = "kernel_balance",
  kernel_balance_centers = 6L,
  kernel_balance_iterations = 400L,
  n_folds = 2L,
  lambda = 0.18,
  washout = 0,
  lag_order = 0,
  rank_max = 3,
  seed = 303
)

dr_balance
```

## Reuse external balancing weights

External weights are useful when the same balancing design should be carried
into the staged fit, DR fit, and post-fit dose-response layer.

```{r}
w_entropy <- cdmc_entropy_balance_weights(
  panel,
  dose = "dose",
  covariates = "x1",
  time = "time",
  time_effects = TRUE
)

dr_external <- cdmc_dr_fit(
  data = panel,
  outcome = "y",
  dose = "dose",
  unit = "unit",
  time = "time",
  covariates = "x1",
  weights = w_entropy,
  n_folds = 2L,
  lambda = 0.18,
  washout = 0,
  lag_order = 0,
  rank_max = 3,
  seed = 303
)

dr_external
```

## Bootstrap DR summaries

```{r}
dr_boot <- cdmc_bootstrap(
  dr_balance,
  n_boot = 10,
  statistics = c("coefficients", "average_tau_dr", "dose_contrast_dr"),
  contrast_history = c(dose_lag0 = 1),
  reference_history = c(dose_lag0 = 0),
  seed = 303
)

dr_boot$summary
```

## Replay alternative weighting scenarios

```{r}
dr_scan <- cdmc_sensitivity_scan(
  dr_balance,
  washout_grid = 0:1,
  weight_specs = list(
    gaussian_linear = list(
      weight_method = "gaussian_gps",
      gps_model = "linear"
    )
  )
)

dr_scan$results[, c("weight_scenario", "fit_success", "weight_mode")]
```

The practical pattern is:

1. use internal Gaussian or kernel GPS weighting when a direct nuisance model
   is acceptable,
2. use balancing-based DR weights when balance itself is the primary design
   target, and
3. use external weights when the same weighting object should propagate across
   several package layers.