---
title: "DR Weighting Workflows"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DR Weighting Workflows}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(causaldosemc)
options(width = 100)
```

This vignette covers the main weighting paths for `cdmc_dr_fit()`: internal
GPS weighting, balancing-based DR fitting, reusable external weights, and DR
replay under alternative nuisance choices.

To keep the focus on weighting rather than support trimming, the examples below
set `washout = 0`.

## Build a panel with richer continuous-dose variation

```{r}
panel <- simulate_cdmc_data(
  n_units = 24,
  n_times = 10,
  beta = 0.85,
  lag_beta = 0.2,
  noise_sd = 0.05,
  switch_on_prob = 0.12,
  switch_off_prob = 0.5,
  seed = 303
)

original_dose <- panel$dose
active <- abs(original_dose) > 0

panel$dose[active] <- sign(original_dose[active]) * (
  0.2 + 0.6 * panel$x1[active] + 0.1 * panel$time[active] / max(panel$time)
)
panel$y <- panel$y + 0.5 * (panel$dose - original_dose)

head(panel)
```

## Fit DR with an internal Gaussian GPS model

```{r}
dr_gps <- cdmc_dr_fit(
  data = panel,
  outcome = "y",
  dose = "dose",
  unit = "unit",
  time = "time",
  covariates = "x1",
  weight_method = "gaussian_gps",
  gps_model = "linear",
  n_folds = 2L,
  lambda = 0.18,
  washout = 0,
  lag_order = 0,
  rank_max = 3,
  seed = 303
)

dr_gps
```

If `mgcv` is installed, the same Gaussian GPS path can use a smoother mean
model for the assignment nuisance layer.

```{r}
if (requireNamespace("mgcv", quietly = TRUE)) {
  dr_gam <- cdmc_dr_fit(
    data = panel,
    outcome = "y",
    dose = "dose",
    unit = "unit",
    time = "time",
    covariates = "x1",
    weight_method = "gaussian_gps",
    gps_model = "gam",
    gps_df = 4L,
    n_folds = 2L,
    lambda = 0.18,
    washout = 0,
    lag_order = 0,
    rank_max = 3,
    seed = 303
  )

  dr_gam
}
```

## Fit DR with internal kernel-balance weights

```{r}
dr_balance <- cdmc_dr_fit(
  data = panel,
  outcome = "y",
  dose = "dose",
  unit = "unit",
  time = "time",
  covariates = "x1",
  weight_method = "kernel_balance",
  kernel_balance_centers = 6L,
  kernel_balance_iterations = 400L,
  n_folds = 2L,
  lambda = 0.18,
  washout = 0,
  lag_order = 0,
  rank_max = 3,
  seed = 303
)

dr_balance
```

## Reuse external balancing weights

External weights are useful when the same balancing design should be carried
into the staged fit, DR fit, and post-fit dose-response layer.

```{r}
w_entropy <- cdmc_entropy_balance_weights(
  panel,
  dose = "dose",
  covariates = "x1",
  time = "time",
  time_effects = TRUE
)

dr_external <- cdmc_dr_fit(
  data = panel,
  outcome = "y",
  dose = "dose",
  unit = "unit",
  time = "time",
  covariates = "x1",
  weights = w_entropy,
  n_folds = 2L,
  lambda = 0.18,
  washout = 0,
  lag_order = 0,
  rank_max = 3,
  seed = 303
)

dr_external
```

## Bootstrap DR summaries

```{r}
dr_boot <- cdmc_bootstrap(
  dr_balance,
  n_boot = 10,
  statistics = c("coefficients", "average_tau_dr", "dose_contrast_dr"),
  contrast_history = c(dose_lag0 = 1),
  reference_history = c(dose_lag0 = 0),
  seed = 303
)

dr_boot$summary
```

## Replay alternative weighting scenarios

```{r}
dr_scan <- cdmc_sensitivity_scan(
  dr_balance,
  washout_grid = 0:1,
  weight_specs = list(
    gaussian_linear = list(
      weight_method = "gaussian_gps",
      gps_model = "linear"
    )
  )
)

dr_scan$results[, c("weight_scenario", "fit_success", "weight_mode")]
```

The practical pattern is:

1. use internal Gaussian or kernel GPS weighting when a direct nuisance model
   is acceptable,
2. use balancing-based DR weights when balance itself is the primary design
   target, and
3. use external weights when the same weighting object should propagate across
   several package layers.

## Choose the orthogonal score and the train baseline weighting

Three orthogonal flags govern the cross-fitted DR machinery. They can be set
independently and are useful for robustness comparisons.

| Flag | Values | What it changes |
|---|---|---|
| `dr_score` | `"aipw"` (default), `"plr"` | The score used to combine the train baseline with the holdout dose history. |
| `baseline_weighting` | `"gps"` (default), `"none"` | Whether the *train-fold* SVT baseline is fit weighted by GPS / balancing weights or unweighted. |
| `weight_method` | `"gaussian_gps"`, `"kernel_gps"`, `"cbps"`, `"entropy_balance"`, `"kernel_balance"`, `"adaptive_balance"`, or external | Which weighting scheme produces the per-cell propensity / balancing weights. |

### When to use each `dr_score`

* `"aipw"` (default, augmented IPW) builds the per-cell pseudo-outcome
  $\tilde\tau_{it} = \hat\tau_{it} + w_{it}\bigl(Y_{it} - \hat Y_{it}^{(0)} - \hat\tau_{it}\bigr)$
  and regresses it on the dose-lag design. It is the historical default; bias
  vanishes if either the baseline model or the GPS weights are correct.
* `"plr"` (Robinson partially linear / Frisch–Waugh) fits a separate
  conditional-mean model $\hat g(X_{it}) = \mathbb{E}[D_{it}\mid X_{it}, t]$ on
  each train fold using the same `gps_model` family, residualizes the holdout
  dose lags as $D_{it} - \hat g(X_{it})$, and regresses outcome residuals on the
  residualized lags **without** GPS reweighting. Consistency does not require
  the GPS weights to be normalised correctly, so `"plr"` is a useful second leg
  for robustness checks.

### When to use each `baseline_weighting`

* `"gps"` (default) reweights the train-fold zero-dose SVT problem by the
  per-cell weights, which can reduce bias when the support of treated cells
  shifts in covariate-treatment-relevant directions.
* `"none"` fits the train-fold baseline unweighted. This is closer to the
  classical AIPW orthogonality argument because the baseline nuisance is
  estimated independently of the propensity nuisance, and it is the fastest
  option. Pair it with a flexible `gps_model` to keep efficiency.

### A robustness contrast in practice

Fit twice with the same data and contrast the dose-lag coefficients:

```{r contrast}
common <- list(
  data = panel, outcome = "y", dose = "dose", unit = "unit", time = "time",
  weight_method = "gaussian_gps", lambda = 0.18, rank_max = 3,
  washout = 0, lag_order = 1, n_folds = 2, seed = 909
)

fit_aipw <- do.call(cdmc_dr_fit, c(common, list(dr_score = "aipw")))
fit_plr  <- do.call(cdmc_dr_fit, c(common, list(dr_score = "plr")))

cdmc_dr_score_contrast(fit_aipw, fit_plr)
```

The returned data frame reports both estimates per dose-lag term, their
absolute and relative differences, and a `sign_agree` flag. Substantial
disagreement, particularly at `dose_lag0`, is a signal that one of the two
nuisances (baseline vs. GPS conditional mean) is poorly specified.

The `summary()` method gives a structured one-pager per fit:

```{r}
summary(fit_aipw)
```
