---
title: "Empirical Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Empirical Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(causaldosemc)
options(width = 100)
```

This vignette shows the package workflow intended for an empirical paper:
blocked-CV main fitting, retuned refit diagnostics, bootstrap with tuning
reruns, and both replay-based and bounded-bias sensitivity summaries.

## Fit the main model with blocked cross-validation

```{r}
panel <- simulate_cdmc_data(
  n_units = 18,
  n_times = 10,
  beta = 0.9,
  lag_beta = 0.3,
  noise_sd = 0.05,
  switch_on_prob = 0.16,
  switch_off_prob = 0.42,
  seed = 202
)

empirical_fit <- cdmc_fit(
  data = panel,
  outcome = "y",
  dose = "dose",
  unit = "unit",
  time = "time",
  covariates = "x1",
  lambda = NULL,
  lambda_selection = "cv",
  cv_rounds = 3L,
  cv_block_size = 2L,
  washout = 1,
  lag_order = 1,
  rank_max = 3,
  seed = 202
)

empirical_fit
empirical_fit$lambda
```

## Run refit diagnostics

The repeated placebo and carryover refit helpers can rerun the same automatic
tuning rule used in the source fit.

```{r}
pretrend <- cdmc_joint_placebo_test(
  empirical_fit,
  placebo_windows = list(
    early_window = -3:-2,
    immediate_window = -1
  ),
  rerun_tuning = TRUE
)

carryover <- cdmc_carryover_refit_test(
  empirical_fit,
  periods = 1:2,
  rerun_tuning = TRUE
)

pretrend
carryover
```

## Screen dynamic assignment restrictions

```{r}
assignment_check <- cdmc_scia_test(
  empirical_fit,
  lags = 2,
  outcome_proxy = "tau",
  restriction_blocks = list(
    recent_history = 1,
    deeper_history = 2
  )
)

assignment_check
```

## Bootstrap the main fit

For substantive work, increase `n_boot` well beyond the small value used here
to keep vignette build times reasonable.

```{r}
empirical_boot <- cdmc_bootstrap(
  empirical_fit,
  n_boot = 10,
  statistics = c("coefficients", "average_tau"),
  rerun_tuning = TRUE,
  seed = 202
)

empirical_boot$summary
```

## Replay washout and support choices

```{r}
empirical_scan <- cdmc_sensitivity_scan(
  empirical_fit,
  washout_grid = 0:2,
  rerun_tuning = TRUE
)

empirical_scan$support_summary
```

## Add formal bounded-bias sensitivity summaries

```{r}
empirical_bounds <- cdmc_sensitivity_bounds(
  empirical_fit,
  statistics = c("coefficients", "average_tau"),
  gamma_grid = c(0, 0.25, 0.5),
  perturbation_layer = "stage2",
  scale = "residual_sd"
)

summary(empirical_bounds)
```

An empirical-paper workflow now has a clear default structure:

1. fit with blocked cross-validation,
2. rerun tuning inside placebo and carryover refits,
3. report bootstrap uncertainty under retuning, and
4. add replay-based and formal sensitivity summaries around the same fit.

For a heavier but closer-to-estimation sensitivity layer, replace
`perturbation_layer = "stage2"` with `perturbation_layer = "optimization"`.