---
title: "Getting Started with causaldosemc"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with causaldosemc}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(causaldosemc)
options(width = 100)
```

This vignette covers the core package entry points for users who want to move
from a simulated panel to a fitted matrix-completion model quickly.

## Simulate a reversible continuous-treatment panel

```{r}
panel <- simulate_cdmc_data(
  n_units = 16,
  n_times = 10,
  beta = 0.8,
  lag_beta = 0.25,
  noise_sd = 0.05,
  switch_on_prob = 0.15,
  switch_off_prob = 0.45,
  seed = 101
)

head(panel)
```

## Fit the default staged estimator

For empirical work, prefer blocked cross-validation when the panel has enough
eligible zero-dose support to tune safely.

```{r}
fit_cv <- cdmc_fit(
  data = panel,
  outcome = "y",
  dose = "dose",
  unit = "unit",
  time = "time",
  covariates = "x1",
  lambda = NULL,
  lambda_selection = "cv",
  cv_rounds = 3L,
  cv_block_size = 2L,
  washout = 1,
  lag_order = 1,
  rank_max = 3,
  seed = 101
)

fit_cv
summary(fit_cv)
head(predict(fit_cv, type = "y0"))
head(predict(fit_cv, type = "tau_model"))
```

## Fit an unbalanced panel

Missing unit-time combinations are padded internally onto the full unit-time
grid and tracked through the `.cdmc_observed` marker.

```{r}
drop_units <- sort(unique(panel$unit))[c(3, 11)]
drop_times <- sort(unique(panel$time))[c(4, 7)]

panel_unbalanced <- subset(
  panel,
  !(unit %in% drop_units & time %in% drop_times)
)

fit_unbalanced <- cdmc_fit(
  data = panel_unbalanced,
  outcome = "y",
  dose = "dose",
  unit = "unit",
  time = "time",
  covariates = "x1",
  lambda = 0.18,
  washout = 1,
  lag_order = 1,
  rank_max = 3,
  seed = 101
)

table(fit_unbalanced$data$.cdmc_observed)
fit_unbalanced
```

## Reuse balancing weights in the main estimator

External weights are useful when the same weighting rule should be reused across
main fitting, DR fitting, and downstream dose-response work.

```{r}
w_entropy <- cdmc_entropy_balance_weights(
  panel,
  dose = "dose",
  covariates = "x1",
  time = "time",
  time_effects = TRUE
)

head(w_entropy$weights)

fit_weighted <- cdmc_fit(
  data = panel,
  outcome = "y",
  dose = "dose",
  unit = "unit",
  time = "time",
  covariates = "x1",
  weights = w_entropy,
  lambda = 0.18,
  washout = 1,
  lag_order = 1,
  rank_max = 3,
  seed = 101
)

fit_weighted
```

## Fit the joint objective

Use the joint objective when a global parametric treatment basis is acceptable
and you want the low-rank baseline and treatment layer solved together.

```{r}
fit_joint <- cdmc_fit(
  data = panel,
  outcome = "y",
  dose = "dose",
  unit = "unit",
  time = "time",
  covariates = "x1",
  objective = "joint",
  effect_model = "linear",
  lambda = 0.18,
  lag_order = 1,
  rank_max = 3,
  seed = 101
)

fit_joint
head(predict(fit_joint, type = "tau_model"))
```

The usual starting point for applied work is still the staged estimator with
blocked cross-validation. The weighted and joint paths are best treated as
deliberate modeling choices rather than automatic upgrades.