---
title: "Manuscript-Style Case Study: Reversible Rebate Intensity"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Manuscript-Style Case Study: Reversible Rebate Intensity}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(causaldosemc)
options(width = 100)
```

This vignette is structured like a compact empirical-paper workflow. The data
are still simulated, but the analysis is written as a single coherent
application rather than a feature-by-feature walkthrough.

The running example treats the continuous, reversible treatment as monthly
rebate intensity across stores. Stores can turn the rebate on and off over
time, rebate intensity varies continuously when active, and effects may persist
for one month after the current rebate choice.

## Build a store-month analysis panel

The starting point is the package simulator, followed by a simple remapping of
the active treatment periods onto a positive rebate-intensity scale. The
outcome is then adjusted so the modified current and lagged rebate levels are
reflected in the observed sales index.

```{r}
panel <- simulate_cdmc_data(
  n_units = 24,
  n_times = 12,
  beta = 0.7,
  lag_beta = 0.25,
  noise_sd = 0.05,
  switch_on_prob = 0.14,
  switch_off_prob = 0.55,
  seed = 505
)

panel <- panel[order(panel$unit, panel$time), ]
original_dose <- panel$dose
active <- abs(original_dose) > 0

panel$dose[active] <- 0.15 +
  0.55 * plogis(panel$x1[active]) +
  0.25 * panel$time[active] / max(panel$time)

lag_original <- ave(original_dose, panel$unit, FUN = function(x) c(0, head(x, -1)))
lag_updated <- ave(panel$dose, panel$unit, FUN = function(x) c(0, head(x, -1)))

panel$y <- panel$y +
  0.7 * (panel$dose - original_dose) +
  0.25 * (lag_updated - lag_original)

case_panel <- data.frame(
  store = panel$unit,
  month = panel$time,
  sales_index = panel$y,
  rebate_intensity = panel$dose,
  demand_proxy = panel$x1
)

head(case_panel)
```

In an empirical application, `sales_index` would be the observed outcome,
`rebate_intensity` would be the store-level treatment intensity, and
`demand_proxy` would stand in for observed demand conditions or local market
controls.

## Fit the primary matrix-completion model

The primary specification uses the staged estimator with blocked
cross-validation, one post-treatment washout period, and a one-lag treatment
effect layer.

```{r}
case_fit <- cdmc_fit(
  data = case_panel,
  outcome = "sales_index",
  dose = "rebate_intensity",
  unit = "store",
  time = "month",
  covariates = "demand_proxy",
  lambda = NULL,
  lambda_selection = "cv",
  cv_rounds = 3L,
  cv_block_size = 2L,
  washout = 1,
  lag_order = 1,
  rank_max = 3,
  seed = 505
)

case_fit
summary(case_fit)
```

This is the baseline empirical specification that the rest of the case study
will stress-test.

## Diagnostic checks before interpretation

The first checks are the repeated placebo refit, the carryover refit, and the
SCIA-style dynamic assignment screen.

```{r}
case_placebo <- cdmc_joint_placebo_test(
  case_fit,
  placebo_windows = list(
    distant_window = -3:-2,
    immediate_window = -1
  ),
  rerun_tuning = TRUE
)

case_carryover <- cdmc_carryover_refit_test(
  case_fit,
  periods = 1:2,
  rerun_tuning = TRUE
)

case_scia <- cdmc_scia_test(
  case_fit,
  lags = 2,
  outcome_proxy = "tau",
  restriction_blocks = list(
    recent_history = 1,
    deeper_history = 2
  )
)

case_placebo
case_carryover
case_scia
```

This is the point in the workflow where an empirical article would decide
whether the baseline identifying story remains plausible enough to move on to
substantive interpretation.

## Add a weighting-oriented robustness fit

Next, the same store-month panel is reweighted with entropy balancing and then
run through the DR layer. This produces a fold-cross-fitted weighted robustness
check using the same treatment and lag structure as the primary analysis.

```{r}
rebate_weights <- cdmc_entropy_balance_weights(
  case_panel,
  dose = "rebate_intensity",
  covariates = "demand_proxy",
  time = "month",
  time_effects = TRUE
)

case_dr <- cdmc_dr_fit(
  data = case_panel,
  outcome = "sales_index",
  dose = "rebate_intensity",
  unit = "store",
  time = "month",
  covariates = "demand_proxy",
  weights = rebate_weights,
  n_folds = 2L,
  lambda = case_fit$lambda,
  washout = 1,
  lag_order = 1,
  rank_max = 3,
  seed = 505
)

case_dr
```

In an article, the point of this section is not to replace the primary fit.
It is to show whether the substantive result looks sensitive to a weighting
design that targets observed dose-covariate balance more aggressively.

## Fit a flexible post-baseline dose-response layer

The main fit anchors the untreated baseline. A flexible response surface can
then be estimated downstream without changing that stage-1 baseline contract.

```{r}
case_curve <- cdmc_dose_response(
  case_fit,
  model = "spline",
  lag_order = 1,
  df = 4
)

predict(
  case_curve,
  dose = seq(0, 1, length.out = 5),
  type = "response"
)
```

## Turn the fitted curve into article-style path contrasts

Two rebate policies are compared against the all-zero path:

- a one-month pulse rebate with high contemporaneous intensity and no lagged
  rebate
- a sustained two-month rebate with moderate intensity in both months

```{r}
case_paths <- cdmc_dynamic_estimand(
  case_curve,
  history = data.frame(
    dose_lag0 = c(0.9, 0.6),
    dose_lag1 = c(0.0, 0.6)
  ),
  reference_history = data.frame(
    dose_lag0 = c(0, 0),
    dose_lag1 = c(0, 0)
  ),
  type = "contrast",
  aggregate = "both",
  labels = c("pulse_rebate", "sustained_rebate")
)

case_paths
case_paths$estimate_table
```

## Bootstrap the path contrasts

The pathwise contrast object is bootstrapable directly, which keeps the paper's
reported uncertainty tied to the exact path definitions used in the text.

```{r}
case_boot <- cdmc_bootstrap(
  case_paths,
  n_boot = 10,
  statistics = "estimate",
  simultaneous = TRUE,
  seed = 505
)

case_boot$summary
```

For a real paper, increase `n_boot` materially beyond the vignette setting.

## Replay washout and weighting sensitivity

The replay-based sensitivity layer is useful for checking whether the primary
result depends sharply on one washout choice or on one weighting decision.

```{r}
case_scan <- cdmc_sensitivity_scan(
  case_fit,
  washout_grid = 0:2,
  weight_specs = list(entropy_reweight = rebate_weights),
  rerun_tuning = TRUE
)

case_scan$support_summary
case_scan$weight_summary
```

## Add formal bounded-bias sensitivity bounds

The bounded-bias layer then asks how large an additive hidden-bias perturbation
would need to be before the main fit's reported summaries lose their sign or
their separation from a chosen null.

```{r}
case_bounds <- cdmc_sensitivity_bounds(
  case_fit,
  statistics = c("coefficients", "average_tau"),
  gamma_grid = c(0, 0.25, 0.5),
  perturbation_layer = "stage2",
  scale = "residual_sd"
)

summary(case_bounds)
```

## What this case study contributes

This vignette is meant to mirror how an empirical paper would actually use the
package:

1. estimate a primary staged matrix-completion model with blocked-CV tuning,
2. run placebo, carryover, and SCIA diagnostics before interpretation,
3. add a weighted DR robustness fit,
4. translate the fitted response surface into policy-relevant dynamic path
   contrasts,
5. bootstrap those reported contrasts, and
6. report both replay-based and bounded-bias sensitivity summaries.

The remaining gap between this workflow and a journal-ready application is now
mostly about real data, empirical reporting choices, and additional substantive
design justification, not about missing package plumbing.